---
title: "Gilteritinib Differntial Protein Analysis"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(amlresistancenetworks)
require(dplyr)
```

## Get Data

This package has formatted the Gilteritnib-treated AML cells into a tidied data frame so they can be easily processed. Here is a summary of the samples collected so that we can better analyze them.

_Samples collected_

```{r Load Data}
gilt.data<-readRDS(system.file('gilteritinibData.Rds',package='amlresistancenetworks'))

#view as table
samps<-gilt.data%>%
  dplyr::select(c(Sample,ligand,CellLine,treatment))%>%distinct()

DT::datatable(samps)
```

## Protein expression

Here we collect the relative expression of each protein across different conditions.

### All Sample Expression

```{r LFC heatmap, echo=FALSE, include=FALSE, warning=FALSE,fig.width=10}

library(pheatmap)

#get data in matrix form
full.mat<-gilt.data%>%
    subset(!is.na(value))%>%
    subset(!is.na(Gene))%>%
    dplyr::select(value,`Sample ID`,Gene)%>%
    tidyr::pivot_wider(names_from='Sample ID',values_from='value',values_fn=list(value=mean))%>%
    tibble::column_to_rownames("Gene")%>%
    as.matrix()

nas<-which(apply(full.mat,1,function(x) any(is.na(x))))
if(length(nas)>0)
  full.mat<-full.mat[-nas,]

vars=apply(full.mat,1,var,na.rm=T)
most.var=names(sort(vars,decreasing=T)[1:50])

full.samps=gilt.data%>%
    dplyr::select(c(`Sample ID`,CellLine,ligand,treatment))%>%
    distinct()%>%tibble::column_to_rownames("Sample ID")

pheatmap(full.mat[most.var,],cellwidth=10,cellheight=10,annotation_col = full.samps, clustering_method = 'ward.D2',file='mostVarProts.png')


```
While there are general patterns between treatment effects and ligands, there are still spurious clusterings brought about by noise making it difficult to draw strong conclustions from the proteins driving these changes.

### Mean condition expression 

Now we can compute the means of the values so that we can see the driving effects. 
```{r mean value heatmap,cho=FALSE, include=FALSE, warning=FALSE,fig.width=10}

mean.vals<-gilt.data%>%
  dplyr::select(CellLine,Gene,treatment,ligand,value)%>%
  group_by(CellLine,Gene,treatment,ligand)%>%
  subset(!is.na(value))%>%
  subset(!is.na(Gene))%>%
  summarize(meanVal=mean(value,na.rm=TRUE))%>%
  rowwise()%>%
  mutate(condition=paste(c(CellLine,treatment,ligand),collapse='_'))%>%
  ungroup()

samps=mean.vals%>%
  dplyr::select(condition,CellLine,treatment,ligand)%>%
  distinct()%>%
  tibble::column_to_rownames('condition')

mat<-mean.vals%>%
  dplyr::select(condition,Gene,meanVal)%>%
  tidyr::pivot_wider(names_from=condition,values_from=meanVal,values_fn=list(meanVal=mean))%>%
  tibble::column_to_rownames('Gene')%>%
  as.matrix()

vars=apply(mat,1,var,na.rm=T)
most.var=names(sort(vars,decreasing=T)[1:50])

pheatmap(mat[most.var,],cellwidth=10,cellheight=10,annotation_col = samps, clustering_method = 'ward.D2',file='mostVarProtMeans.png')

```

For the purposes of this study, we are interested in determining _which_ proteins are driving changes in drug response, not necessarily how similar/different they are. For this we compute the mean differences between protein activity for specific conditions of interest.

## Differential expression

We compute a number of differential expression calculations. For each differential expression calculation we do the following:
* Identify the number of differentially expressed proteins
* Identify any enriched/depleted pathways via GSEA

```{r raw differences,echo=FALSE, include=FALSE, warning=FALSE,fig.width=10}

#rcalculate differences and p-values

early.data<-gilt.data%>%
    subset(treatment%in%(c('None','Early Gilteritinib')))%>%
    dplyr::select(Gene,Sample,CellLine,ligand,value)%>%
    rename(treatment='ligand')

total.mean.diffs<-amlresistancenetworks::computeFoldChangePvals(early.data,control='None',conditions=c("FL","FGF2"))
molm.mean.diffs<-amlresistancenetworks::computeFoldChangePvals(subset(early.data, CellLine=='MOLM14'),
                                                          control='None',
                                                          conditions=c("FL","FGF2"))
mv411.mean.diffs<-amlresistancenetworks::computeFoldChangePvals(subset(early.data, CellLine=='MV411'),
                                                          control='None',
                                                          conditions=c("FL","FGF2"))

mean.diffs<-rbind(mutate(total.mean.diffs,CellLine='Both'),
                  mutate(molm.mean.diffs,CellLine='MOLM14'),
                  mutate(mv411.mean.diffs,CellLine='MV411'))

#count the proteins at our significance threshold
prot.counts=mean.diffs%>%
    subset(p_adj<0.05)%>%
    group_by(Condition,CellLine)%>%
    summarize(`ProteinsDiffEx`=n_distinct(Gene))


prots=total.mean.diffs%>%
    subset(p_adj<0.05)%>%ungroup()%>%
  dplyr::select(Gene)

pheatmap(full.mat[intersect(rownames(full.mat),as.character(prots$Gene)),],cellwidth=10,cellheight=10,annotation_col = full.samps, clustering_method = 'ward.D2',file='diffExProts.png')

DT::datatable(prot.counts)


```
Now we can see that the differential proteins seem to cluster beter, and could provide interesting biomarkers.



### FGF2 treatment in both cell lines

In both cell lines we evaluate the GO term enrichment using the `clusterProfiler` package along side the `MSigDB` `C2` gene lists. We use an FDR p-value correction and limit the genes searched to those in the protein universe - i.e. those that were measured in the initial dataset.


```{r gene lists,echo=FALSE,warning=FALSE}

#first we break down the gene lists
mv411<-subset(mean.diffs,CellLine=='MV411')
molm14<-subset(mean.diffs,CellLine=='MOLM14')
tots<-subset(mean.diffs,CellLine=='Both')

prot.univ<-unique(gilt.data$Gene)

```

#### FGF2 treated MV411, MOLM14 and both

FGF2-treate cells show up-regulation of a handful of pathways including immune-related pathways.

```{r,echo=FALSE, warning=FALSE,fig.width=10}
genes.with.values=mv411%>%
    ungroup()%>%
    subset(Condition=='FGF2')%>%
    dplyr::select(Gene,value=condition_to_control)

mv411.fgf2=computeGSEA(genes.with.values,prot.univ)

enrichplot::ridgeplot(mv411.fgf2)+ggplot2::ggtitle("GO Terms for MV411 FGF2")
  ggplot2::ggsave('MV411_FGF2_GO.png',width=16,height=8)

genes.with.values=molm14%>%
    ungroup()%>%
    subset(Condition=='FGF2')%>%
    dplyr::select(Gene,value=condition_to_control)

molm14.fgf2=computeGSEA(genes.with.values,prot.univ)

enrichplot::ridgeplot(molm14.fgf2)+ggplot2::ggtitle("GO Terms for MOLM14 FGF2")
  ggplot2::ggsave('molm14_FGF2_GO.png',width=16,height=8)

genes.with.values=tots%>%
    ungroup()%>%
    subset(Condition=='FGF2')%>%
    dplyr::select(Gene,value=condition_to_control)

tot.fgf2=computeGSEA(genes.with.values,prot.univ)

enrichplot::ridgeplot(tot.fgf2)+ggplot2::ggtitle("GO Terms for Combined FGF2")
  ggplot2::ggsave('total_FGF2_GO.png',width=16,height=8)

  DT::datatable(subset(mean.diffs,p_adj<0.05))

```




### FL treatment in both cell lines

Now let's measure the impact of FLT3 ligand in both cell lines. It seems to up-regulate similar pathwaays


```{r,echo=FALSE, warning=FALSE,fig.width=10}

genes.with.values=mv411%>%
    ungroup()%>%
    subset(Condition=='FL')%>%
    dplyr::select(Gene,value=condition_to_control)

mv411.fl=computeGSEA(genes.with.values,prot.univ)


if(nrow(as.data.frame(mv411.fl))>0){
  enrichplot::ridgeplot(mv411.fl,showCategory=20)+ggplot2::ggtitle("GO Terms for MV411 FLT3 Ligand")
  ggplot2::ggsave('MV411_FL_GO.png',width=16,height=8)
}


genes.with.values=molm14%>%
    ungroup()%>%
    subset(Condition=='FL')%>%
    dplyr::select(Gene,value=condition_to_control)

molm14.fl=computeGSEA(genes.with.values,prot.univ)

if(nrow(as.data.frame(molm14.fl))>0){
  enrichplot::ridgeplot(molm14.fl,showCategory=20)+ggplot2::ggtitle("GO Terms for MOLM14 FLT3 Ligand")
  ggplot2::ggsave('MOLM14_FL_GO.png',width=16,height=8)
}



genes.with.values=tots%>%
    ungroup()%>%
    subset(Condition=='FL')%>%
    dplyr::select(Gene,value=condition_to_control)

tot.fl=computeGSEA(genes.with.values,prot.univ)

if(nrow(as.data.frame(tot.fl))>0){
  enrichplot::ridgeplot(tot.fl,showCategory=20)+ggplot2::ggtitle("GO Terms for both cells FLT3 Ligand")
  ggplot2::ggsave('total_FL_GO.png',width=16,height=8)
}


```

We have similar pathways enriched.

### FGF2 vs FL treatment

So our primary question is realy to identify what is similar and different between the two ligand treatments. For that we originally just take the difference of differences between the two. Those proteins that are highly similar in FC or different in FC will show up at alternate ends of the list and can therefore be compared.

 
```{r diff, echo=FALSE,warning=FALSE,fig.width=10}

mean.diffs<-amlresistancenetworks::computeFoldChangePvals(early.data,
                                                          control='FL',
                                                          conditions=c("FGF2"))

mv411.diffs<-amlresistancenetworks::computeFoldChangePvals(subset(early.data, CellLine=='MV411'),
                                                          control='FL',
                                                          conditions=c("FGF2"))

molm14.diffs<-amlresistancenetworks::computeFoldChangePvals(subset(early.data,CellLine=='MOLM14'),
                                                          control='FL',
                                                          conditions=c("FGF2"))

tot.diff<-mean.diffs%>%dplyr::select(Gene,value=condition_to_control)
  

combined.diff=computeGSEA(tot.diff,prot.univ)

if(nrow(as.data.frame(combined.diff))>0){
  enrichplot::ridgeplot(combined.diff,showCategory=20)+ggplot2::ggtitle("GO Terms for all FGF2 vs FLT3 Ligase")
    ggplot2::ggsave('bothCells_FL_vs_FGF2_GO.png',width=16,height=8)

}

DT::datatable(subset(mean.diffs,p_adj<0.05))

##now compute for MV411
tot.diff<-mv411.diffs%>%dplyr::select(Gene,value=condition_to_control)
  

combined.diff=computeGSEA(tot.diff,prot.univ)

if(nrow(as.data.frame(combined.diff))>0){
  enrichplot::ridgeplot(combined.diff,showCategory=20)+ggplot2::ggtitle("GO Terms for MV411 FGF2 vs FLT3 Ligase")
    ggplot2::ggsave('MV411_FL_vs_FGF2_GO.png',width=16,height=8)

}

DT::datatable(subset(mv411.diffs,p_adj<0.05))

##now compute for MOLM
tot.diff<-molm14.diffs%>%dplyr::select(Gene,value=condition_to_control)
  

combined.diff=computeGSEA(tot.diff,prot.univ)

if(nrow(as.data.frame(combined.diff))>0){
  enrichplot::ridgeplot(combined.diff,showCategory=20)+ggplot2::ggtitle("GO Terms for MOLM FGF2 vs FLT3 Ligase")
    ggplot2::ggsave('MOLM_FL_vs_FGF2_GO.png',width=16,height=8)

}
DT::datatable(subset(molm14.diffs,p_adj<0.05))


```
