---
title: "Figure 5 Best cell line signatures"
author: "Sara Gosline"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(amlresistancenetworks)

if(!exists('dataLoaded')){
  amlresistancenetworks::loadBeatAMLData()
  dataLoaded=TRUE
}

##assumes figure 2 was already generated!!!!
reg.preds<-readRDS('cellLinelassoRegPreds.rds')%>%
  mutate(compound=as.character(compound),Molecular=as.character(Molecular))
log.preds<-readRDS('cellLinelogRegPreds.rds')%>%
  mutate(compound=as.character(compound),Molecular=as.character(Molecular))


syn=synapseLogin()

## get depmap
cells<-syn$tableQuery('select distinct Sample from  syn23004114 where ( LogFoldChange<>0 AND transcriptCounts<>0)')$asDataFrame()$Sample

all.cells<-syn$tableQuery('select distinct Sample from syn23674921 where ( LogFoldChange<>0 AND transcriptCounts<>0)')$asDataFrame()$Sample

#cl.mol.dat<-syn$tableQuery('select * from syn23674921')$asDataFrame()%>%
  #syn23004114')$asDataFrame()%>%
#  subset(Sample%in%all.cells)%>%
#
cl.mol.dat<-syn$tableQuery('select * from syn23004114')$asDataFrame()%>%
  subset(Sample%in%cells)%>%
  dplyr::rename(mRNALevels='transcriptCounts',proteinLevels='LogFoldChange',
                geneMutations='numMuts')%>%
  mutate(Sample=unlist(Sample),Gene=unlist(Gene))%>%
    rowwise()%>%
  mutate(binaryMutations=ifelse(geneMutations==0,0,1))%>%
        mutate(binaryMutations=tidyr::replace_na(binaryMutations,0))


#create manual mapping from AML to cell Line
drug.mapping <- syn$tableQuery('select * from syn23193914 where PTRC is not NULL')$asDataFrame()

##now match the drugs between cl and tumors
cl.auc.dat<-syn$tableQuery('select * from syn23680639')$asDataFrame()%>%
  subset(Metric=='auc')%>%
  subset(Condition%in%drug.mapping$drugName)%>%
  mutate(drugName=unlist(Condition),Sample=unlist(Sample))%>%
  left_join(drug.mapping,by=c('drugName','source'))%>%
  dplyr::select(Sample,Condition='PTRC',Value,source)%>%
  subset(Sample%in%cl.mol.dat$Sample)%>%
  group_by(Sample,Condition,source)%>%summarize(Value=mean(Value))%>%ungroup()

sanger.cl.auc<-cl.auc.dat%>%
  subset(source=='Sanger')%>%
  mutate(AUC=(Value*100+50))

ctrp.cl.auc<-cl.auc.dat%>%
  subset(source=='CTRP')%>%
  mutate(AUC=Value*10)

```

## Best signatures from LASSO

From the predictions we can select those results that have the _best_ response, and plot those below. 

```{r plotting functions}
#these functions just enable plotting on cell line data
selectCellDataMatAndPlot<-function(compound,method,Molecular,
                                   genelist, doEnrich, dataset){
  
  #print(paste(c(compound,method,Molecular)))
  
  dat.mat<-dplyr::rename(cl.mol.dat,value=Molecular)
  if(dataset=='CTRP') 
    c.auc.dat<- dplyr::select(ctrp.cl.auc,-c(Value,source))
  else
    c.auc.dat<- dplyr::select(sanger.cl.auc,-c(Value,source))
  
    
  genes=getFeaturesFromString(genelist,Molecular)

  clusterSingleDrugEfficacy(drugName=compound,
                            meth=method,
                            data=Molecular,
                            doEnrich=doEnrich,
                            auc.dat=c.auc.dat,
                            auc.thresh=100,
                            genes=genes,
                            data.mat=dat.mat,
                            prefix=dataset)  
    

}
```

Next we can compare the *best*-performing signatures on the cell line data.

```{r lasso preds,warning=FALSE,echo=FALSE}
  new.results<-reg.preds%>%
  subset(testMSE<1800)%>%  
    mutate(method='LASSO',doEnrich=TRUE)%>%
    dplyr::rename(genelist='genes')%>%
   rowwise()%>%
    mutate(enrich=paste(selectCellDataMatAndPlot(compound,method,Molecular,genelist,doEnrich,dataset)$Description,collapse=','))%>%
    rowwise()%>%
    mutate(features=paste(getFeaturesFromString(genelist,Molecular),collapse=','))%>%
    dplyr::select(dataset,Drug='compound',testMSE,features,dataType='Molecular',enrich)%>%
    arrange(testMSE)


 knitr::kable(new.results)
```

## best signatures for logistic regression

Here we use another threshold, that of 0.3, to evaluate the best signatures using a logistic regression.

```{r logr preds, echo=FALSE, warning=FALSE}

  other.results<-log.preds%>%subset(numFeatures>2)%>%subset(testMSE<0.45)%>%  
    mutate(method='LogisticRegression',doEnrich=TRUE)%>%
    dplyr::rename(genelist='genes')%>%
   rowwise()%>%
   mutate(enrich=paste(selectCellDataMatAndPlot(compound,method,Molecular,genelist,doEnrich,dataset)$Description,collapse=','))%>%
    rowwise()%>%
    mutate(features=paste(getFeaturesFromString(genelist,Molecular),collapse=','))%>%
    dplyr::select(dataset,Drug='compound',testMSE,features,dataType='Molecular',enrich)%>%
  arrange(testMSE)


knitr::kable(other.results)
full.results<-rbind(mutate(new.results,Model='LASSO'),
                    mutate(other.results,Model='Logistic'))

write.csv(full.results,'table5_bestCellLineSigs.csv')

```

Now just look at just Trametinib and Quizartinib

Moving forward we want to see which results are most robust, so turn to the cell line data
```{r lasso results}
 lasso.results<-reg.preds%>%subset(numFeatures>2)%>%
  subset(compound%in%c('Trametinib (GSK1120212)','Quizartinib (AC220)'))%>%
  #subset(testMSE<1800)%>%  
    mutate(method='LASSO',doEnrich=TRUE)%>%
    dplyr::rename(genelist='genes')%>%
   rowwise()%>%
    mutate(enrich=paste(selectCellDataMatAndPlot(compound,method,Molecular,genelist,doEnrich,dataset)$Description,collapse=','))%>%
    rowwise()%>%
    mutate(features=paste(getFeaturesFromString(genelist,Molecular),collapse=','))%>%
    dplyr::select(dataset,Drug='compound',testMSE,features,dataType='Molecular',enrich)%>%
    arrange(testMSE)

  knitr::kable(lasso.results)
```
LASSO results are not great, maybe logistic is better. 
```{r logistic results}
 log.results<-log.preds%>%subset(numFeatures>2)%>%
  subset(compound%in%c('Trametinib (GSK1120212)','Quizartinib (AC220)'))%>%
  #subset(testMSE<1800)%>%  
    mutate(method='Logistic',doEnrich=TRUE)%>%
    dplyr::rename(genelist='genes')%>%
   rowwise()%>%
    mutate(enrich=paste(selectCellDataMatAndPlot(compound,method,Molecular,genelist,doEnrich,dataset)$Description,collapse=','))%>%
    rowwise()%>%
    mutate(features=paste(getFeaturesFromString(genelist,Molecular),collapse=','))%>%
    dplyr::select(dataset,Drug='compound',testMSE,features,dataType='Molecular',enrich)%>%
    arrange(testMSE)

  knitr::kable(log.results)
```

